Efficient pedigree recording for fast population genetics simulation

### 2017-11-20 forwards_paper.tex
removed trailing spaces
fix citations

\jda{Is this the right place for this paragraph? should it maybe be integrated
	with graf beginning 'Modern computing' above?}
\plr{this paragraph is supposed to cite the fwds time simulation methods we need to - @molpopgen fix this up?}
Forwards-time population genetics simulations are growing in importance,
as the genome-wide impacts of hitchhiking and background selection are more widely appreciated.
For instance, \citet{harris2016genetic} used SLiM \citep{haller2017flexible} to simulate tens of thousands of human exomes
to assess the impact of genetic load and Neanderthal introgression on human genetic diversity.
This is roughly state-of-the-art 
-- the fastest simulation frameworks such as SLiM and fwdpp \citep{fwdpp}
can ``only'' simulate tens of megabases of sequence in tens of thousands of individuals
for tens of thousands of generations.
This is still substantially short of whole genomes (hundreds to thousands of megabases)
for many biological population sizes of interest 
(e.g., millions for a restricted population of mosquitos).
Achieving these sizes will likely require parallelization \citep{lawrie2017accelerating}.
The general idea of storing genealogical information to speed up simulations is not new --
it was entirely implemented in AnA-FiTS \citep{aberer2013rapid}, for instance --
but this program did not implement the critical step of discarding irrelevant genealogical information.
A similar but more limited method for discarding this information does appear in \citet{padhukasahasram2008exploring}.

@@

The \msprime{} tables API is independent of the details of a particular simulation engine.  In other words, one simply
needs to fill the relevant columns and send them to \msprime{} at regular intervals for simplification.  For example,
using the \fwdpp{} C++ API \cite{Thornton2014-hx}, we fill arrays of simple structures representing the node and edge data.
Doing so requires no modification of the \fwdpp{} code base.  Rather, we simply need to book-keep the parent/offspring
labels and perform simple processing of the recombination breakpoints from each mating event.  Using \texttt{pybind11}
(\url{https://github.com/pybind/pybind11/}),
the node/edge arrays on
the C++ side are made visible, without copy, to Python as a multi-column numpy array. This array is in turn processed by
the \msprime{} tables API.  In order to create a standalone user interface, our implementation uses \texttt{pybind11} to
create a Python package extending \texttt{fwdpy11} (\url{https://github.com/molpopgen/fwdpy11}), which is a Python
package based on \fwdpp{}.  (A neat aside, that we may wish to skip here, is that it may be possible to use the tables
API from a command-line C/C++ application via an embedded Python interpreter.)
\plr{@molpopgen: note on how this works with fwdpp?}
\krt{@plr: done, but perhaps too verbose?}

>>>>

\jda{Is this the right place for this paragraph? should it maybe be integrated
	with graf beginning 'Modern computing' above?}
\plr{this paragraph is supposed to cite the fwds time simulation methods we need to - @molpopgen fix this up?}
Forwards-time population genetics simulations are growing in importance,
as the genome-wide impacts of hitchhiking and background selection are more widely appreciated.
For instance, \citet{harris2016genetic} used SLiM \citep{slim`} to simulate tens of thousands of human exomes
to assess the impact of genetic load and Neanderthal introgression on human genetic diversity.
This is roughly state-of-the-art
-- the fastest simulation frameworks such as SLiM and fwdpp \citep{fwdpp}
can ``only'' simulate tens of megabases of sequence in tens of thousands of individuals
for tens of thousands of generations.
This is still substantially short of whole genomes (hundreds to thousands of megabases)
for many biological population sizes of interest
(e.g., millions for a restricted population of mosquitos).
Achieving these sizes will likely require parallelization \citep{lawrie2017accelerating}.
The general idea of storing genealogical information to speed up simulations is not new --
it was entirely implemented in AnA-FiTS \citep{aberer2013rapid}, for instance --
but this program did not implement the critical step of discarding irrelevant genealogical information.
A similar but more limited method for discarding this information does appear in \citet{padhukasahasram2008exploring}.

@@

The \msprime{} tables API is independent of the details of a particular simulation engine.  In other words, one simply
needs to fill the relevant columns and send them to \msprime{} at regular intervals for simplification.  For example,
using the \fwdpp{} C++ API \cite{fwdpp}, we fill arrays of simple structures representing the node and edge data.
Doing so requires no modification of the \fwdpp{} code base.  Rather, we simply need to book-keep the parent/offspring
labels and perform simple processing of the recombination breakpoints from each mating event.  Using \texttt{pybind11}
(\url{https://github.com/pybind/pybind11/}),
the node/edge arrays on
the C++ side are made visible, without copy, to Python as a multi-column numpy array. This array is in turn processed by
the \msprime{} tables API.  In order to create a standalone user interface, our implementation uses \texttt{pybind11} to
create a Python package extending \texttt{fwdpy11} (\url{https://github.com/molpopgen/fwdpy11}), which is a Python
package based on \fwdpp{}.  (A neat aside, that we may wish to skip here, is that it may be possible to use the tables
API from a command-line C/C++ application via an embedded Python interpreter.)
\plr{@molpopgen: note on how this works with fwdpp?}
\krt{@plr: done, but perhaps too verbose?}
